---
title: "Civil servants by sexual orientation (by profession and department, 31 March 2020)"
author: "Andi Fugard"
date: "26/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get the data

It's linked [over here](https://www.gov.uk/government/statistics/civil-servants-by-sex-and-ethnicity-and-disability-and-sexual-orientation-by-profession-and-department-2020).

```{r message=FALSE}
library(readxl)
library(tidyverse)
localExcelTemp <- tempfile(fileext = ".xlsx")
download.file("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/938871/Civil_servants_by_sex_and_ethnicity_and_disability_and_sexual_orientation_by_profession_and_department__2020.xlsx",
              destfile = localExcelTemp,
              mode = "wb")
messy_sheet <- read_xlsx(localExcelTemp,
                         "Sexual Orientation",
                         skip = 4) %>% na.omit()
names(messy_sheet)[1] <- "dept"
```

```{r}
head(messy_sheet)
```

## Tidy is a bit

First record the dept names that have numbers due to footnote superscripts:

```{r}
messy_sheet <- messy_sheet %>%
  mutate(dept = recode(dept, `Defence Electronics & Components Agency 7` = "Defence Electronics & Components Agency",
         `Submarine Delivery Agency 8` = "Submarine Delivery Agency",
         `Forestry and Land Scotland 9` = "Forestry and Land Scotland",
         `Scottish Forestry 10` = "Scottish Forestry"))
```



### Save the margins

Using `as.numeric` to zap anything which doesn't look like a number.

```{r}
dept_margins <- messy_sheet %>%
  select(dept, Total) %>%
  mutate(perc = as.numeric(Total)*100,
         Total = NULL)
dept_margins
```



```{r}
job_margins <- messy_sheet %>%
  filter(dept == "All employees") %>%
  select(-c(dept, Total)) %>%
  t() %>%
  as.data.frame() %>%
  rename(prop = V1) %>%
  mutate(job = rownames(.),
         perc = as.numeric(prop) * 100,
         prop = NULL)
job_margins
```

### Tidy up the rest

Remove the margin totals in the spreadsheet and pivot longer:

```{r}
tidy_sheet <- messy_sheet %>%
  filter(dept != "All employees") %>%
  select(-Total) %>%
  pivot_longer(cols = !"dept", names_to = "job", values_to = "prop")
head(tidy_sheet)
```

Remove all the missingness, whatever the reason, and transform the proportions in E-notation to percentages.

```{r}
tidier_sheet <- tidy_sheet %>%
  filter(!prop %in% c("..", "--", "-", "*")) %>%
  mutate(perc = as.numeric(prop)*100) %>%
  select(-prop)
head(tidier_sheet)
```


## Make some pictures


### Departments

```{r fig.height=12, fig.width=7}
dept_margins %>%
  na.omit() %>%
  ggplot(aes(x = reorder(dept, perc), y = perc)) +
  geom_point(aes(colour = dept), size = 2) +
  coord_flip() +
  labs(x = NULL, y = "%",
       title = "LGBO civil servants") +
  ylim(0,20) +
  theme(legend.position = "none")
```



### Professions

```{r fig.height=6, fig.width=6}
job_margins %>%
  na.omit() %>%
  ggplot(aes(x = reorder(job, perc), y = perc)) +
  geom_point(aes(colour = job)) +
  coord_flip() +
  labs(x = NULL, y = "%",
       title = "LGBO civil servants") +
  ylim(0,20) +
  theme(legend.position = "none")
```


### Look within depts and professions

Let's choose a few jobs from the list. The bigger the blob, the higher the percentage in that profession and dept.

```{r fig.height=6, fig.width=9}
tidier_sheet %>%
  filter(job %in% c("Social Research", "Economics", "Statistics", "Policy")) %>%
  ggplot(aes(x = job, y = reorder(dept, desc(dept)), colour = dept)) +
  geom_point(aes(size = perc)) +
  theme(legend.position = "none") +
  labs(x = "Profession",
       y = NULL,
       title = "LGBO civil servants",
       subtitle = "Blob size proportional to %",
       caption = "There are many reasons for missing blobs - see original spreadsheet") +
  scale_x_discrete(position = "top") 
```





